---
title: "Assignment 3"
author: "James Lopez"
date: "October 11, 2016"
output: pdf_document
---
##2 Write a simulation to calculate the power for the following study design. The study has two variables, treatment group and outcome. There are two treatment groups (0, 1) and they should be assigned randomly with equal probability. The outcome should be a random normal variable with a mean of 60 and standard deviation of 20. If a patient is in the treatment group, add 5 to the outcome. 5 is the true treatment effect. Create a linear model for the outcome by the treatment group, and extract the p-value (hint: see assigment1). Test if the p-value is less than or equal to the alpha level, which should be set to 0.05. Repeat this procedure 1000 times. The power is calculated by finding the percentage of times the p-value is less than or equal to the alpha level. Use the set.seed command so that the professor can reproduce your results.

#part 1 Find the power when the sample size is 100 patients.
```{r}
set.seed(100)

loops <- 1e3
exp.pvals <- numeric(loops)
for(i in seq(loops)) {
  group <- rnorm(100, 60, 20)
  assign <- sample(c(0,1), 100, replace = TRUE)
  exp <- data.frame(group, assign)
  for(x in 1:100) {
    if(exp[x,'assign'] == 1) {
      exp[x,1] <- exp[x,1] +5
    }
    x <- x+1
  }
  mod.sim <- lm(group~assign, data = exp)
  summ <- summary(mod.sim)
  coefs <- coef(summ)
  exp.pvals[i] <- coefs[2,4]
  i <- i+1
}

sum(exp.pvals<=0.05)/loops
```

The power for 100 samples is 0.231.

#part 2 Find the power when the sample size is 1000 patients.

```{r}
set.seed(1000)
loops_2 <- 1e3
exp.pvals_2 <- numeric(loops_2)
for(j in seq(loops_2)) {
  group_2 <- rnorm(1000, 60, 20)
  assign_2 <- sample(c(0,1), 1000, replace = TRUE)
  exp_2 <- data.frame(group_2, assign_2)
  for(y in 1:1000) {
    if(exp_2[y,'assign_2'] == 1) {
      exp_2[y,1] <- exp_2[y,1] +5
    }

  }
  mod.sim_2 <- lm(group_2~assign_2, data = exp_2)
  summ_2 <- summary(mod.sim_2)
  coefs_2 <- coef(summ_2)
  exp.pvals_2[j] <- coefs_2[2,4]

}

sum(exp.pvals_2<=0.05)/loops_2
```

The power for 1000 sampes is 0.98.



##3 Obtain a copy of the football-values lecture. Save the 2016/proj_wr16.csv file in your working directory. Read in the data set and remove the first two columns.

The football-values data set is stored in data frame footbal, which is then converted into football with the first two columns of the original data set removed.

```{r}
footbal <- read.csv('proj_wr16.csv')
football <- footbal[,-1:-2]
```

#part 1 Show the correlation matrix of this data set.

The correlation matrix of football is shown using:
```{r}
cor(football)
```

#part 2 Generate a data set with 30 rows that has a similar correlation structure. Repeat the procedure 10,000 times and return the mean correlation matrix.

To make a data set with a similar correlation structure, we use mvrnorm since the football data set consists of multiple variables and corresponding means and covariances. The resulting correlation matrix of the simulated data is contained in keep.i.

```{r}
library(MASS)
rho <- cor(football)
vcov <- var(football)
means <- colMeans(football)

keep.i <- 0
loops <- 1e4
for(i in seq(loops)) {
  football.sim <- mvrnorm(30, mu=means, Sigma=vcov)
  keep.i <- keep.i + cor(football.sim)/loops
}

keep.i
```

#part 3 Generate a data set with 30 rows that has the exact correlation structure as the original data set.

The procedure for making a simulated data set with the SAME correlation structure as the football data set is the same, except we include the argument 'empirical' and set it equal to TRUE to show that the means and variances we use to create our simulated data are empirical means and variances rather than true means and variances.

```{r}
keep.j <- 0
loops <- 1e4
for(j in seq(loops)) {
  football.sim <- mvrnorm(30, mu=means, Sigma=vcov, empirical = TRUE)
  keep.j <- keep.j + cor(football.sim)/loops
}
keep.j
```


##Question 4: \LaTeX expressions 

#part 1


\begin{equation}
P(B) = \sum_j P(B | A_j)P(A_j), 
\Rightarrow P(A_i | B) = \frac {P(B | A_i)P(A_i)}{\sum_j P(B | A_j)P(A_j)}
\end{equation}


#part 2


\begin{equation}
\hat{f}(\zeta) = \int_{-\infty}^{\infty} f(x)e^{-2\pi i x \zeta}dx
\end{equation}


#part 3


\begin{equation}
\textbf{J} = \frac {d\textbf{f}}{d\textbf{x}} = \bigg[\frac{\partial\textbf{f}}{\partial x_1} \cdots \frac{\partial\textbf{f}}{\partial x_n}\bigg] = \begin{bmatrix}
\frac{\partial f_1}{\partial x_1} & \cdots & \frac{\partial f_1}{\partial x_n} \\
\vdots & \ddots & \vdots \\
\frac{\partial f_m}{\partial x_1} & \cdots & \frac{\partial f_m}{\partial x_n}
\end{bmatrix}
\end{equation}


