---
title: "Assignment 2"
author: "James Lopez"
date: "October 11, 2016"
output: pdf_document
---

**Grade: 48/50**

##1. 
#part i.
```{r}
#setwd("~/Bios 6301")
cancer.df <- read.csv("cancer.csv", header = TRUE)
```

#part ii.
```{r}
nrow(cancer.df)
ncol(cancer.df)
# There are 42120 rows and 8 columns in cancer.df
```

#part iii.
```{r}
colnames(cancer.df)
#The column names of cancer.df are "year", "site", "state", "sex", "race", "mortality", "incidence", and "population"
```

#part iv.
```{r}
cancer.df[3000,6]
# The value in the 3000th row and 6th column in cancer.df is 350.69
```

#part v.
```{r}
cancer.df[172,]
```

#part vi.
```{r}
cancer.df[,'incidence rate'] <- cancer.df[,'incidence']/100000
```

**JC Grading - 1**
For incidence rate above should be incidence / population * 100000

#part vii.
```{r}
sum(cancer.df[,'incidence rate'] == 0)
#There are 23191 subgroups with a zero incidence rate
```

#part viii.
```{r}
which.max(cancer.df[,'incidence rate'])
cancer.df[21387,]
#The subgroup with the highest incidence rate is in row 21387 of cancer.df, during the year 2002 in California for white females with breast cancer.
```

**JC Grading - 1**
syntax is fine but answer is incorrect b/c of how incidence rate was calculated


##2. 
#part i.

The sum() function should produce an error because it only works with numeric data types. The quotations around the numbers in x assigns the numbers as characters instead of numerics.
```{r}
x <- c("5","12","7")
max(x) 
sort(x) 
#sum(x) **Not included as an R statement in rmd due to compiling issues
```

The max is "7" because "12" begins with a 1 and "5" begins with a 5. Since x is a character vector, the "numbers" will be treated like characters and be sorted by the first character in the whole string. "Max" will not produce an error.

Because x has numbers as characters, the value of 12 has the smallest value because it starts with a 1, while 7 has the largest value because it starts with a 7. So "Sort" will not produce an error.



#part ii.

The next two commands will produce an error message because the elements in vector y do not have the same data type. The numeric elements 7 and 12 will be coerced into characters due to the presence of "5"
Because the vector will coerce the 7 and 12 into characters, the + operator will not work.
```{r}
#y <- c("5",7,12)
#y[2] + y[3]

# NOTE: Not included as R statements due to compiling issues
```

#part iii.
```{r}
z <- data.frame(z1="5",z2=7,z3=12)
z[1,2] + z[1,3]
```

The + operator returned a number 19. 
z was made into a data frame, which allows the elements of z to retain their original data type.
z[1,2] and z[1,3] called for the numeric elements in z, so the + operator was able to be used to return numeric 19.


##3.

#part i.
```{r}
c(1:8, 7:1)
```

#part ii.
```{r}
rep(1:5, 1:5)
```

#part iii.
```{r}
m <- matrix(1, nrow = 3, ncol = 3)
for(i in 1:3) {
  m[i,i] <- 0
}
m
```

#part iv.
```{r}
n <- matrix(c(1:4), nrow = 4, ncol = 4, byrow = TRUE)
for (i in 1:4) {
  n[i,] <- n[i,]^i
}
n
```


##4.

#part i.

I made a function that tried to follow the equation posted. To demonstrate the function, I randomly selected both the base and the exponential order between 1-10.

```{r}
h <- function(x,n){
n <- sample.int(10,1)
x <- sample.int(10,1)
for(i in 1:n) {
  print(sum(x^(1:i)))
}
}
h(x,n)
```

#part ii. a.
```{r}
j <- 1
for (i in 1:999) {
  if (i %% 3 == 0 | i %% 5 == 0) {
    j <- j+i
  }
}
j
```

#part ii. b.
```{r}
k <- 1
for (i in 1:999999) {
  if (i %% 4 == 0 | i %% 7 == 0) {
    k <- k+i
  }
}
k
```

**JC Grading +0 Bonus**

